% Thesis Published Work ---------------------------------------------------
\chapter{Publications}

Publications arising in the course of this thesis:

\itemize

\item{Read H., Johnson S., Barquist L., Mills G., Gardner P.P., Patrick W.M., Wiles S. \textbf{The effect of constitutive bioluminescence expression on the in vitro and in vivo fitness of the mouse enteropathogen \textit{Citrobacter rodentium}.} Manuscript in preparation.}

\item{Wong V., Pickard D., Barquist L., Sivaraman K., Harte P., Arends M., Kane L., Mottram L., Ellison L., Kay S., Wileman T., MacLennan, Kingsley R.A., Dougan G.  \textbf{Characterization of the yehUT two-component regulatory system of \textit{Salmonella enterica} serovars Typhi and Typhimurium}. Manuscript in preparation.}

\item{Okoro C.K., Barquist L., Kingsley R.A., Connor T.R., Harris S.R., Arends M., Stevens M., Parry C.M., Al-Mashhadani M.N., Kariuki S., Msefula C.L., Gordon M.A., de Pinna E., Wain J., Heyderman R.S., Obaro S., Alonso P.L., Mandomando I., MacLennan C.A., Tapia M.D., Levine M.M., Tennant S.M., Parkhill J., Dougan G.  \textbf{Signatures of adaptation in human invasive \textit{S.} Typhimurium populations.} Manuscript in preparation.} 

\item{Wilf N.M., Reid A.J., Ramsay J.P., Williamson N.R., Croucher N.J., Gatto L., Hester S.S., Goulding D., Barquist L., Lilley K.S., Kingsley R.A., Dougan G., Salmond G.P.C.. \textbf{RNA-seq reveals the RNA binding proteins, Hfq and RsmA, play various roles in virulence, antibiotic production and genomic flux in \textit{Serratia} sp. 39006}. Manuscript under review.}

\newpage

\item{Pettit L.J., Browne H.P., Yu L., Smits W.K., Fagan R.P., Barquist L., Martin M.J., Goulding D., Duncan S.H., Flint H.J., Dougan G., Choudhary J.S., Lawley T.D. \textbf{Functional genomics reveals that \textit{Clostridium difficile} Spo0A coordinates sproulation, virulence and metabolism.} Manuscript under review.}

\item{Reuter S., Connor T.R., Barquist L., Walker D., Feltwell T., Harris S.R., Fookes M., Hall M.E., Fuchs T.M., Corander J., Dufour M., Ringwood T., Savin C., Bouchier C., Martin L., Miettinen M., Shubin M., Laukkanen-Ninios R., Sihvonen L.M., Siitonen A., Skurnik M., Falc\~{a}o J.P., Fukushima H., Scholz H.C., Prentice M., Wren B.W., Parkhill J., Carniel E., Achtman M., McNally A., Thomson N.R. \textbf{Parallel independent evolution of pathogenicity within the genus \textit{Yersinia}.} Manuscript under review.}

\item{Barquist L., Burge S.W., Gardner P.P. \textbf{Building non-coding RNA families.} \textit{Methods in Molecular Biology}, in press.} 

\item{Hoeppner M.P., Barquist L., Gardner P.P. \textbf{An introduction to RNA databases.} \textit{Methods in Molecular Biology}, in press.} 

\item{Croucher N.J., Mitchell A.M., Gould K.A., Inverarity D., Barquist L., Feltwell T., Fookes M.C., Harris S.R., Dordel J., Salter S.J., Browall S., Zemlickova H., Parkhill J., Normark S., Henriques-Normark B., Hinds J., Mitchell T.J., Bentley S.D. \textbf{Dominant role of nucleotide substitution in the diversification of serotype 3 pneumococci over decades and during a single infection.} \textit{PLoS Genetics}, 2013.} 

\item{Kingsley R.A., Whitehead S., Connor T., Barquist L., Sait L., Holt K., Sivaraman K., Wileman T., Goulding D., Clare S., Hale C., Seshasayee A., Harris S., Thomson N., Gardner P., Rabsch W., Wigley P., Humphrey T., Parkhill J., Dougan G. \textbf{Genome and transcriptome adaptation accompanying emergence of the DT2 host-restricted \textit{Salmonella} Typhimurium pathovar.}  \textit{mBio}, 2013.}

\item{Martin M.J., Clare S., Goulding D., Faulds-Pain A., Barquist L., Browne H., Pettit L., Dougan G., Lawley T.D.,  Wren B.W. \textbf{The \textit{agr} locus regulates virulence and colonization genes in \textit{Clostridium difficile} 027.} \textit{Journal of Bacteriology}, 2013. }

\newpage

\item{Barquist L., Boinett C.J., Cain A.K..  \textbf{Approaches to querying bacterial genomes with transposon-insertion sequencing.} \textit{RNA Biology}, 10(7), 2013.}

\item{Barquist L., Langridge G.C., Turner D.J., Phan M.D., Turner A.K., Bateman A., Parkhill J., Wain J., Gardner P.P. \textbf{A comparison of dense transposon insertion libraries in the \textit{Salmonella} serovars Typhi and Typhimurium.} \textit{Nucleic Acids Research}, 41(8):4549-4564, 2013.} 

\item{Burge S.W., Daub J., Eberhardt R., Tate J., Barquist L., Nawrocki E.P., Eddy S.R., Gardner P.P., Bateman A. \textbf{Rfam 11.0: 10 years of RNA families.} \textit{Nucleic Acids Research}. 41(D1):D226-D232, 2013}

\item{Croucher N.J., Harris S.R., Barquist L., Parkhill J., Bentley S.D. \textbf{A high-resolution view of genome-wide pneumococcal transformation.} \textit{PLoS Pathogens}, 8(6), 2012}

\item{Westesson O., Barquist L., Holmes I. \textbf{HandAlign: Bayesian multiple sequence alignment, phylogeny, and ancestral reconstruction.} \textit{Bioinformatics}, 28(8):1170-1171, 2012}

\item{Gardner P.P., Barquist L., Bateman A., Nawrocki E.P., Weinberg Z. \textbf{RNIE: genome-wide prediction of bacterial intrinsic terminators.} \textit{Nucleic Acids Research}, 39(14):5845-5852, 2011}



 
 % ----------------------------------------------------------------------

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End: 
