% Thesis Acknowledgements ------------------------------------------------


\begin{acknowledgementslong} 
%uncommenting this line, gives a different acknowledgements heading
%\begin{acknowledgements}      %this creates the heading for the acknowlegments


This PhD project is a result of great ideas, hard work and unconditional support of numerous people that I would hereby like to acknowledge. 

Firstly, I would like to thank my supervisor Anton Enright for the opportunity to conduct my PhD project in his research group. He provided support and patience, while allowing me in the same time to follow my own course of studies. Most importantly, through his group he provided a great working environment, both scientifically engaging and fun. 

I offer my sincerest gratitude to the two leaders of my main collaborative projects, Donal O'Carroll and Duncan Odom. They provided me with all the attention, help and resources as if I was their group member. Their experience, guidance, and most importantly their passion have greatly improved my work. Donal, thank you for your clarity and determination, it kept me motivated when I was doing my first steps in the field. Duncan, thank you for your energy, for weeks of manuscript writing long in the night at the Orchard street, as well as for the guidance for my next steps in science. 

%I express my gratitude to my Thesis Committee: Eileen Furlong, Alvis Brazma and Eric Miska. They reviewed each step of my PhD studies and provided valuable suggestions for my projects. 

A big thanks goes to the members of the Enright laboratory. Cei Abreu-Goodger  was my unofficial mentor in the first year and taught me the importance of scrutiny and the systematic approach. Mat Davis remains the best resource of novel ideas and sanity checks, in combination with his British calmness and politeness. My gratitude also goes to the other members of the lab: Stijn van Dongen, Harpreet Kaur Saini, Afonso Guerra Assuncao, Leonor Quintais and Sergei Manakov from which I have learned a lot during the last years.

I would also like to thank PhD students and postdocs from the laboratories of my collaborators that performed the experimental part of the research. I will always admire Serena De Fazio for her strength and perseverance. My great gratitude goes to Claudia Kutter, with whom it was an immense pleasure to work with. The collaboration was a book example of the exchange of ideas, mutual motivation and critical discussions that improved the quality of work at each step. Thank you for making the last part of my PhD a brilliant experience.

This thesis would not see the light of day without all the corrections and reading from Anton Enright, Mat Davis, Myrto Kostadima, Mikhail Spivakov, Konrad Rudolph, Felix Krueger, Tom Rensch and Christine Seeliger. I thank you greatly for your time and patience and I will use your advice in the future.

I am indebted to my teachers and professors from Croatia and the United World College USA. Darinka Cvetko and Katica Sekulic started my love for biology and mathematics. Mirela Kurnik and Romana Halapir Frankovic showed me that mathematicians and biologists are fun people doing awesome things. Fernando Mejia, Kurt Lucas and Judy Land taught me how to think outside the boundaries of the literature. Kristian Vlahovicek lead me into the field of bioinformatics itself. All of them invested a great deal of time and effort to transfer their hard learned knowledge and gave me a chance, for which I am incredibly grateful.

Furthermore, I would like to thank all my friends that kept me sane throughout the PhD years and were my Cambridge family. First and foremost Angela Goncalves and Petra Schwalie with whom I started the EMBL PhD and shared most of my Cambridge experiences. I also thank the people that marked my first formative Cambridge year: Marija Buljan, Tina Perica, Andre Faure, Sergei Manakov, Leopold Parts, Matias Piipari and Steve Pettit. I would also like to thank my closest ones: Felix Krueger, Myrto Kostadima, Benedetta Baldi, Michele Mattioni, Sander Timmer, Christine Seeliger, Benjamin Stauch, Mikhail Spivakov and Angela Goncalves for being there, making the mayhem and making it all worth it. Also have to mention Heidi, Julia, Judith, Micha, Konrad, Jean-Baptiste, Steve, Nils and Remco.  I will miss the long breakfasts, beers at the Elm tree and the mad travelling. A big thank you to Pembroke rowers, especially Chris and Chris. And I will make sure that Iva, Marin, Marwan and Damir keep close. Hopefully no friendship was harmed during the making of this thesis. Though a birthday or two might have been forgotten. I have learnt from all of you so much more than from my PhD work alone. 

Special thanks to my housemates, Christine and Sander, who kept me going with their wisdom and their wit. They always increased my mood and occasionally my weight. Wiggle mansion forever.

Finally, I would like to thank my family for the amount of love and unconditional support that one can only hope for. 

%\end{acknowledgements}
\end{acknowledgementslong}

% ------------------------------------------------------------------------

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End: 
